Phenotypic evolution is more constrained in simpler food webs

### 2019-01-17 manuscript.tex
updated MS prior to sending to Chris, Brendan and Arezoo. I don't like the fact that manuscript creates a cache of figures that aren't in the manuscript, but are in the reproduce_analyses cache too. Need to fix this.
rerun reproducible analyses for manuscript

\documentclass[11pt,]{article}
\usepackage{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{natbib}
\bibliographystyle{amnat}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Matthew A. Barbour\^{}\{1,2,\textbackslash{}ast\}; Christopher J. Greyson-Gaito\^{}\{1,3\}; Arezoo Sootodeh\^{}\{3\}; Brendan Locke\^{}\{4\}; Jordi Bascompte\^{}\{2\}},
            pdftitle={Phenotypic evolution is more constrained in simpler food webs},
            colorlinks=true,
            citecolor=black,
            urlcolor=black,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

  \title{Phenotypic evolution is more constrained in simpler food webs}
    \author{Matthew A. Barbour\(^{1,2,\ast}\) \\ Christopher J. Greyson-Gaito\(^{1,3}\) \\ Arezoo Sootodeh\(^{3}\) \\ Brendan Locke\(^{4}\) \\ Jordi Bascompte\(^{2}\)}
    \date{}

%%% From AmNat_MS_template.tex
%\documentclass[11pt]{article}
%\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
%\usepackage[authoryear,sectionbib,sort]{natbib}
\linespread{1.7}
%\usepackage[utf8]{inputenc}
\usepackage{lineno}

%%%%%%% Sections not currently working properly
%\usepackage{titlesec}
%\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
%\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
%\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
%\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}

%%% My added packages
%\usepackage{amsmath} % for matrices
%\usepackage{graphicx} % for organizing figures
%\usepackage{lmodern} % for tilde
%\usepackage[T1]{fontenc} % for tilde

%%%%%%%%% Previous - likely useful though
%\usepackage{setspace}\doublespacing
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

@@

\maketitle

\noindent 1. University of British Columbia, Department of Zoology,
Vancouver, BC V6T 1Z4, Canada;

\noindent 2. University of Zurich, Department of Evolutionary Biology
and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich,
Switzerland;

@@

\emph{Manuscript elements}: Figure 1, figure 2, figure 3, figure 4. All
figures should be printed in color.

@@

\emph{Keywords}: fitness landscape, host-parasitoid, natural selection,
G-matrix, community context, ecological networks, evolutionary ecology,
eco-evolutionary dynamics.

@@

The adaptive landscape provides a powerful framework for understanding
how natural selection has shaped the evolution of biodiversity ---from
genes to phenotypes to species
\citep{Wright1931, Simpson1944, Arnold2001}. More than a metaphor, the
adaptive landscape links quantitative genetic and phenotypic variation
to evolution by natural selection
\citep{Lande1979, Arnold1984applications, Arnold1984theory}. Ecological
interactions often play a key role in shaping adaptive landscapes, as
evidenced by the role of antagonistic \citep{Schluter2000, Abrams2000}
and mutualistic \citep{Bronstein2006} interactions in driving
evolutionary change. Although there is clear evidence that pairwise
interactions can shape the adaptive landscape, we do not know how the
adaptive landscape is shaped by the community context
\citep{McPeek2017, terHorst2018}. Resolution on how the community
context shapes phenotypic evolution is urgently needed though, given the
rapid impacts of climate change on ecological communities
\citep{Scheffers2016}.

@@

We collected data on three different traits that we expected to
influence larval survival based on our previous work \citep{Barbour2016}
and others work with gall midges \citep{Weis1983, Heath2018}. First, we
measured gall diameter as the size of each gall chamber to the nearest
0.01 mm at its maximum diameter (perpendicular to the direction of plant
tissue growth). Our previous work has shown that larger galls are
associated with higher survival \citep{Barbour2016}. Second, we measured
clutch size by counting the number of chambers in each gall
\citep{Weis1983, Heath2018}. All larvae collected from the same
multi-chambered gall were scored with the same clutch size. Third, we
measured oviposition preference as the density of larvae observed on a
plant in an independent survey. We did this by randomly sampling five
branches per tree and counting the number of individual gall chambers
(number of larvae). We then converted these counts to a measure of
larval density per 100 shoots by counting the number of shoots on the
last branch we sampled. All larvae collected from the same plant were
scored with the same oviposition preference. Measuring larval densities
on plants in the field is a common method for measuring oviposition
preference \citep{Gripenberg2010}; however, caution must be taken in
inferring `preference' as larval densities can be influenced by
processes other than preference \citep{Singer1986}. Fortunately, a
couple of features of our study system suggest that larval densities may
be a good proxy for oviposition preference. For example, since our data
comes from a randomized placement of plant genotypes in a common garden,
there is no consistent bias in which plant genotypes females are exposed
to while searching for oviposition sites. Also, egg predation is a minor
source of mortality for galling insects in general \citep{Hawkins1997};
therefore, we do not expect any prior egg predation to bias our
estimates of observed larval densities.

@@

Rather than imposing selection, parasitoids may influence the expression
of herbivore traits which could bias measurements of selection. In our
system, it was plausible that parasitoids may influence chamber diameter
by altering larval feeding behavior or killing larvae before they
complete their development. To estimate this potential bias, we subset
our data to only include galls where there was variation in larval
survival within the same gall (i.e.~1 \textgreater{} mean survival
\textgreater{} 0). If we assume that larvae within each gall should have
similar chamber diameters because they come from the same clutch and
experience the same local environment (an assumption our data supports:
gall ID explains 54\% of the variance in chamber diameter), then the
relationship between chamber diameter and larval survival in this data
subset represents the effect of parasitism on trait expression
(i.e.~bias). We used a GLMM with the same structure as described above
except that we only modelled a linear relationship between chamber
diameter and larval survival (\(\alpha_{\text{Diam}}\)). We detected a
positive bias in both food-web treatments (complex
\(\alpha_{\text{Diam}}\)= 0.36 {[}0.05, 0.67{]}; simple
\(\alpha_{\text{Diam}}\)= 0.42 {[}0.01, 0.82{]}), indicating that
unadjusted relationships would overestimate the strength of selection on
chamber diameter. To account for this bias, we substracted our mean
estimates of bias from our estimates with the full dataset prior to
calculating chamber diameter's trait-fitness relationship and
directional selection gradient.

@@

\indent We found that more gall midge traits experienced directional
selection in the simple (3 of 3) vs.~complex food-web treatment (1 of
3)(table \ref{Table:Gradients}). For example, we observed directional
selection for smaller clutch sizes in the simple food web, but there was
no evidence of selection acting on this trait in the complex food web
(table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}C). This
absence of selection appeared to be a result of conflicting selection
pressures imposed by each guild of parasitoids (fig.
\ref{fig:UV_Landscape}). Specifically, when we subset our data to focus
on differences between parasitoid guilds, we found that larval
parasitoids actually impose directional selection for larger clutch
sizes (\(\beta_{\text{Clutch}}\)= 0.13 {[}0.03, 0.23{]}). In the simple
food web, we also observed clear evidence of directional selection for
midges to avoid ovipositing on plants with high densities of
conspecifics (table \ref{Table:Gradients}; fig.
\ref{fig:UV_Landscape}B); however, this relationship was weaker in
complex food webs (table \ref{Table:Gradients}). This was likely a
result of larval parasitoids imposing greater mortality on egg
parasitoids at high gall midge densities (see Selection on the extended
phenotype of egg parasitoids section), and thus a less than additive
effect on gall midges. Chamber diameter experienced positive directional
selection in both food-web treatments, but selection was 65\% higher in
complex food webs (table \ref{Table:Gradients}; fig.
\ref{fig:UV_Landscape}). This was not due to any difference between egg
and larval parasitoids (fig. \ref{fig:UV_Landscape}), but was due to the
lower average survival in complex food webs (contrast -0.26 {[}-0.12,
-0.42{]}).

@@

\begin{table}[h]
\caption{Standardized selection gradients acting on gall midges in complex vs. simple food webs.}
\label{Table:Gradients}
\centering
\begin{tabular}{lccc}
\\ 
\hline
\textbf{Selection gradient} & \textbf{Complex} & \textbf{Simple} & \textbf{Contrast}  \\ 
\hline
$\beta_{\text{Diam}}$ & 
\textbf{
0.34 [
0.22,
0.47] }& 
\textbf{
0.21 [
0.13,
0.3] }&

-0.13 [
-0.26,
0.01] \\

$\beta_{\text{Clutch}}$ & 
0.06 [
-0.04,
0.17] & 
\textbf{
-0.09 [
-0.18,
-0.02] }& 
\textbf{
-0.15 [
-0.29,
-0.02] }\\

$\beta_{\text{Pref}}$ &
-0.12 [
-0.29,
0.05] & 
\textbf{
-0.16 [
-0.25,
-0.07] }&

-0.04 [
-0.22,
0.14] \\

$\gamma_{\text{Diam,Diam}}$ &
0.14 [
-0.06,
0.32] &

0.1 [
-0.03,
0.25] &

-0.04 [
-0.27,
0.2] \\

$\gamma_{\text{Clutch,Clutch}}$ & 
-0.05 [
-0.25,
0.16] &

-0.12 [
-0.27,
0.03] &

-0.06 [
-0.33,
0.19] \\

$\gamma_{\text{Pref,Pref}}$ & 
\textbf{
0.33 [
0.06,
0.64] }&

0.01 [
-0.14,
0.17] & 
\textbf{
-0.32 [
-0.66,
0] }\\

$\gamma_{\text{Diam,Clutch}}$ & 
-0.05 [
-0.17,
0.08] &

@@

-0.02 [
-0.16,
0.13] \\

$\gamma_{\text{Diam,Pref}}$ & 
-0.13 [
-0.28,
0] &

-0.02 [
-0.09,
0.06] &

0.12 [
-0.05,
0.29] \\

$\gamma_{\text{Clutch,Pref}}$ & 
0.03 [
-0.11,
0.18] &

@@

-0.03 [
-0.19,
0.12] \\ 
\hline
\end{tabular}
\bigskip{}
\\
{\footnotesize Note: Values in brackets represent 95\% confidence intervals. Bold values indicate that the 95\% CI does not overlap zero. $\beta_{\text{Diam}}$ has been adjusted for bias.}
\end{table}

@@

\indent The curvature of the adaptive landscape indirectly affects
genetic constraints and is influenced by directional, quadratic, and
correlational selection gradients. Our food-web treatment did not alter
correlational selection for any combination of traits (table
\ref{Table:Gradients}). Similarly, our food-web treatment did not
influence quadratic selection on either chamber diameter or clutch size
(table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}A,C). In
contrast, our food-web treatment did alter quadratic selection acting on
oviposition preference (table \ref{Table:Gradients}). The negative
relationship between oviposition preference and larval survival dampened
at high densities in complex, but not simple food webs (fig.
\ref{fig:UV_Landscape}B). This dampened relationship was partly due to a
trend for nonlinear selection imposed by larval parasitoids
(\(\gamma_{\text{Pref:Pref}}\)= 0.18 {[}-0.01, 0.42{]}), but was also
magnified by the lower average survival in complex food webs.

@@

\[\textbf{C}_{\text{Complex}} = \begin{pmatrix} 
-0.12 &  &  \\  
0 & 0 &  \\  
0 & 0 & 0.33 \end{pmatrix}\]

@@

Gall midge preference influenced the probability of observing egg
parasitoids in complex vs.~simple food webs (contrast
\(\alpha_{\text{Pref}}\)= 1 {[}0.03, 2.22{]} ). Specifically, larval
parasitoids decreased the probability of observing egg parasitoids at
high gall midge densities (fig. \ref{fig:EggPtoid_Selection}). Note
however that this trait-fitness relationship did not translate into a
significant difference in selection (contrast \(\beta_{\text{Pref}}\)=
0.24 {[}-0.09, 0.62{]} ). This was likely because there was a tendency
(albeit not statistically signficant) for a lower probability of
egg-parasitoid survival in complex vs.~simple food webs (contrast=0.1
{[} -0.08, 0.31{]}), which masks this trait-fitness relationship. We
also observed two other interesting trends, although neither
significantly differed between treatments. Specifically, there was a
tendency for the impact of larval parasitoids to increase nonlinearly at
higher gall midge densities (fig. \ref{fig:EggPtoid_Selection}; complex
\(\alpha_{\text{Pref:Pref}}\)= -0.49 {[}-1.08, -0.1{]}; but contrast
\(\alpha_{\text{Pref:Pref}}\)= 0.46 {[}-0.03, 1.1{]}). There was also a
tendency for egg parasitoids to obtain a refuge in large gall chambers
at high gall midge densities (complex \(\alpha_{\text{Pref:Diam}}\)=
0.25 {[}0.02, 0.5{]}; but contrast \(\alpha_{\text{Pref:Diam}}\)= -0.25
{[}-0.57, 0.04{]}).

@@

The generality of our results likely depends on the relative abundance
and functional differences between consumers in a community. In the
simplest case, when consumers do not differ from each other, then we do
not expect changes in food-web complexity to modify selective
constraints. Also, many consumers may be at too low of abundances to
impose selection on their resources. Rank abundance curves
\citep{Preston1948} and the disproportionate number of weak interactions
in diverse communities \citep{Paine1992} support this notion. When
consumers are abundant though, the effect of food-web complexity will
depend on whether different species impose conflicting selection
pressures or select for distinct traits. For example, parisitoids and
birds impose conflicting selection pressures on the size of galls
induced by the fly \emph{Eurosta solidaginis}
\citep{Weis1985, Abrahamson1997, Start2016}. Recent studies in this
system have shown that decreases in the relative abundance of birds, due
to either small patch sizes \citep{Start2016} or proximity to urban
areas \citep{Start2018} causes a shift from neutral to directional
selection on gall size. On the other hand, different consumers may
impose selection on different traits, favoring trait integration and
increasing genetic constraints. Examples of this include strong genetic
covariances in plant resistance to different insect herbivores
\citep{Maddox1990, Wise2007, Wise2013}, although there are also examples
where these covariances are weak \citep{Roche1997, Barbour2015}, or vary
from year-to-year \citep{Johnson2007}. We suggest that gaining
predictive insight to the effects of food-web disassembly requires an
understanding of the mechanisms governing the assembly of trophic
interactions.

Our results suggest that the simplification of ecological communities
may not only directly affect connected species, but also result in
indirect evolutionary effects. In our study, this indirect effect arises
from egg parasitoids being released from intraguild predation in simpler
food webs. This release occurs more on trees with high larval densities,
which could intensify selection on gall midge oviposition preference. A
growing number of experiments over the past two decades have
demonstrated the presence and potential importance of indirect
evolutionary effects that emerge in a community context
\citep{Juenger1998, Stinchcombe2001, Lankau2007, Walsh2008, Walsh2010, terHorst2010, Sahli2011, Lau2012, terHorst2015, Schiestl2018}.
If indirect evolutionary effects are common \citep{Walsh2013}, then
predicting evolutionary trajectories resulting from the simplification
of food webs will require evolutionary studies to explicitly account for
the ecological networks that species are embedded in.

@@

\section{Acknowledgements}\label{acknowledgements}

We thank the staff of Humboldt Bay National Wildlife Refuge (U.S. Fish
and Wildlife Service) for facilitating experimental logistics. For
assistance with fieldwork, we thank Ruthie Espanol and Andrew MacDonald.
For funding support, we thank the University of British Columbia (James
Robert Thompson Fellowship and Four-Year Fellowship to M.A.~Barbour),
NSERC (Discovery grant to Greg Crutsinger), and the Swiss National
Science Foundation (grant 31003A\_160671 to J. Bascompte).

\section{Tables}\label{tables}

>>>>

\documentclass[11pt,]{article}
\usepackage[]{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=black,
            urlcolor=black,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{amnat}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{}
    \pretitle{\vspace{\droptitle}}
  \posttitle{}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

%%% Modified from Latex Template for Am. Nat., many other details aren't necessary because they are specified in YAML header of Rmarkdown document.
\usepackage{fullpage}
\linespread{1.7}
\usepackage{lineno}

% code below is important for holding figure positions in main text. Make sure to set fig.pos="H" in code chunk for figure.
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

@@

\vspace*{0.1cm}

\begin{center} \LARGE Phenotypic evolution is more constrained in simpler food webs \end{center}

\begin{center} \large Matthew A. Barbour$^{1,2,\ast}$, Christopher J. Greyson-Gaito$^{2,3}$, Arezoo Sootodeh$^{2}$, Brendan Locke$^{4}$, Jordi Bascompte$^{1}$ \normalsize \end{center}

\noindent 1. University of Zurich, Department of Evolutionary Biology
and Environmental Studies, Zurich, 8057 ZH, Switzerland;

\noindent 2. University of British Columbia, Department of Zoology,
Vancouver, BC V6T 1Z4, Canada;

@@

\emph{Manuscript elements}: table 1, figure 1, figure 2, figure 3,
figure 4. All figures should be printed in color.

@@

\emph{Keywords}: adaptive landscape, ecological networks,
eco-evolutionary dynamics, natural selection, host-parasitoid.

\emph{Number of words}: 5075

@@

The adaptive landscape provides a powerful framework for understanding
how natural selection has shaped the evolution of biodiversity ---from
genes to phenotypes to species
\citep{Wright1931, Simpson1944, Arnold2001}. More than a metaphor, the
adaptive landscape links quantitative genetic and phenotypic variation
to evolution by natural selection
\citep{Lande1979, Arnold1984applications, Arnold1984theory}. Ecological
interactions often play a key role in shaping adaptive landscapes, as
evidenced by the role of antagonistic \citep{Schluter2000, Abrams2000}
and mutualistic \citep{Bronstein2006} interactions in driving
evolutionary change. Although there is clear evidence that pairwise
interactions can shape the adaptive landscape, we still have a poor
understanding of how the adaptive landscape is shaped by the community
context \citep{McPeek2017, terHorst2018}. Resolution on how the
community context shapes phenotypic evolution is urgently needed though,
given the rapid impacts of climate change on ecological communities
\citep{Scheffers2016}.

@@

We collected data on three different traits that we expected to
influence larval survival based on our previous work \citep{Barbour2016}
and other work with gall midges \citep{Weis1983, Heath2018}. First, we
measured gall diameter as the size of each gall chamber to the nearest
0.01 mm at its maximum diameter (perpendicular to the direction of plant
tissue growth). Our previous work has shown that larger galls are
associated with higher survival \citep{Barbour2016}. Second, we measured
clutch size by counting the number of chambers in each gall
\citep{Weis1983, Heath2018}. All larvae collected from the same
multi-chambered gall were scored with the same clutch size. Third, we
measured oviposition preference as the density of larvae observed on a
plant in an independent survey. We did this by randomly sampling five
branches per tree and counting the number of individual gall chambers
(number of larvae). We then converted these counts to a measure of
larval density per 100 shoots by counting the number of shoots on the
last branch we sampled. All larvae collected from the same plant were
scored with the same oviposition preference. Measuring larval densities
on plants in the field is a common method for measuring oviposition
preference \citep{Gripenberg2010}; however, caution must be taken in
inferring `preference' as larval densities can be influenced by
processes other than preference \citep{Singer1986}. Fortunately, a
couple of features of our study system suggest that larval densities may
be a good proxy for oviposition preference. For example, since our data
comes from a randomized placement of plant genotypes in a common garden,
there is no consistent bias in which plant genotypes females are exposed
to while searching for oviposition sites. Also, egg predation is a minor
source of mortality for galling insects in general \citep{Hawkins1997};
therefore, we do not expect any prior egg predation to bias our
estimates of observed larval densities.

@@

Rather than imposing selection, parasitoids may influence the expression
of herbivore traits which could bias measurements of selection. In our
system, it was plausible that parasitoids may influence chamber diameter
by altering larval feeding behavior or killing larvae before they
complete their development. To estimate this potential bias, we subset
our data to only include galls where there was variation in larval
survival within the same gall (i.e.~1 \textgreater{} mean survival
\textgreater{} 0). If we assume that larvae within each gall should have
similar chamber diameters because they come from the same clutch and
experience the same local environment (an assumption our data supports:
gall ID explains 54\% of the variance in chamber diameter), then the
relationship between chamber diameter and larval survival in this data
subset represents the effect of parasitism on trait expression
(i.e.~bias). We used a GLMM with the same structure as described above
except that we modelled only a linear relationship between chamber
diameter and larval survival (\(\alpha_{\text{Diam}}\)). We detected a
positive bias in both food-web treatments (complex
\(\alpha_{\text{Diam}}\)= 0.36 {[}0.05, 0.67{]}; simple
\(\alpha_{\text{Diam}}\)= 0.42 {[}0.01, 0.82{]}), indicating that
unadjusted relationships would overestimate the strength of selection on
chamber diameter. To account for this bias, we substracted our mean
estimates of bias from our estimates with the full dataset prior to
calculating chamber diameter's trait-fitness relationship and
directional selection gradient.

@@

\indent We found that more gall midge traits experienced directional
selection in the simple (3 of 3) vs.~complex food-web treatment (1 of
3)(table \ref{Table:Gradients}). For example, we observed directional
selection for smaller clutch sizes in the simple food web, but there was
no evidence of selection acting on this trait in the complex food web
(table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}C). This
absence of selection appeared to be a result of conflicting selection
pressures imposed by each guild of parasitoids (fig.
\ref{fig:UV_Landscape}). Specifically, when we subset our data to focus
on differences between parasitoid guilds, we found that larval
parasitoids actually impose directional selection for larger clutch
sizes (\(\beta_{\text{Clutch}}\)= 0.13 {[}0.04, 0.24{]}). In the simple
food web, we also observed clear evidence of directional selection for
midges to avoid ovipositing on plants with high densities of
conspecifics (table \ref{Table:Gradients}; fig.
\ref{fig:UV_Landscape}B); however, this relationship was weaker in
complex food webs (table \ref{Table:Gradients}). This was likely a
result of larval parasitoids imposing greater mortality on egg
parasitoids at high gall midge densities (see Selection on the extended
phenotype of egg parasitoids section), and thus a less than additive
effect on gall midges. Chamber diameter experienced positive directional
selection in both food-web treatments, but selection was 66\% higher in
complex food webs (table \ref{Table:Gradients}; fig.
\ref{fig:UV_Landscape}). This was not due to any difference between egg
and larval parasitoids (fig. \ref{fig:UV_Landscape}), but was due to the
lower average survival in complex food webs (contrast -0.27 {[}-0.11,
-0.42{]}).

@@

\begin{table}[h]
\caption{Standardized selection gradients acting on gall midges in complex vs. simple food webs.}
\label{Table:Gradients}
\centering
\begin{tabular}{lccc}
\\ 
\hline
\textbf{Selection gradient} & \textbf{Complex} & \textbf{Simple} & \textbf{Contrast}  \\ 
\hline
$\beta_{\text{Diam}}$ & 
\textbf{
0.34 [
0.22,
0.48] }& 
\textbf{
0.21 [
0.12,
0.31] }&

-0.14 [
-0.27,
0] \\

$\beta_{\text{Clutch}}$ & 
0.06 [
-0.05,
0.17] & 
\textbf{
-0.09 [
-0.17,
-0.01] }& 
\textbf{
-0.15 [
-0.29,
-0.03] }\\

$\beta_{\text{Pref}}$ &
-0.13 [
-0.29,
0.05] & 
\textbf{
-0.16 [
-0.26,
-0.06] }&

-0.03 [
-0.21,
0.15] \\

$\gamma_{\text{Diam,Diam}}$ &
0.13 [
-0.06,
0.33] &

0.1 [
-0.02,
0.23] &

-0.03 [
-0.27,
0.2] \\

$\gamma_{\text{Clutch,Clutch}}$ & 
-0.05 [
-0.27,
0.18] &

-0.11 [
-0.28,
0.03] &

-0.06 [
-0.32,
0.2] \\

$\gamma_{\text{Pref,Pref}}$ & 
\textbf{
0.34 [
0.07,
0.63] }&

0.02 [
-0.14,
0.18] & 
\textbf{
-0.32 [
-0.64,
-0.01] }\\

$\gamma_{\text{Diam,Clutch}}$ & 
-0.04 [
-0.16,
0.08] &

@@

-0.02 [
-0.17,
0.12] \\

$\gamma_{\text{Diam,Pref}}$ & 
-0.13 [
-0.29,
0.02] &

-0.02 [
-0.1,
0.07] &

0.12 [
-0.05,
0.3] \\

$\gamma_{\text{Clutch,Pref}}$ & 
0.03 [
-0.1,
0.18] &

@@

-0.03 [
-0.18,
0.12] \\ 
\hline
\end{tabular}
\bigskip{}
\\
{\footnotesize Note: Values in brackets represent 95\% confidence intervals. Bold values indicate that the 95\% CI does not overlap zero. $\beta_{\text{Diam}}$ has been adjusted for bias.}
\end{table}

@@

\indent The curvature of the adaptive landscape indirectly affects
genetic constraints and is influenced by directional, quadratic, and
correlational selection gradients. Our food-web treatment did not alter
correlational selection for any combination of traits (table
\ref{Table:Gradients}). Similarly, our food-web treatment did not
influence quadratic selection on either chamber diameter or clutch size
(table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}A,C). In
contrast, our food-web treatment did alter quadratic selection acting on
oviposition preference (table \ref{Table:Gradients}). The negative
relationship between oviposition preference and larval survival dampened
at high densities in complex, but not simple food webs (fig.
\ref{fig:UV_Landscape}B). This dampened relationship was partly due to a
trend for nonlinear selection imposed by larval parasitoids
(\(\gamma_{\text{Pref:Pref}}\)= 0.18 {[}-0.02, 0.42{]}), but was also
magnified by the lower average survival in complex food webs.

@@

\[\textbf{C}_{\text{Complex}} = \begin{pmatrix} 
-0.12 &  &  \\  
0 & 0 &  \\  
0 & 0 & 0.34 \end{pmatrix}\]

@@

Gall midge preference influenced the probability of observing egg
parasitoids in complex vs.~simple food webs (contrast
\(\alpha_{\text{Pref}}\)= 1.03 {[}0.04, 2.31{]} ). Specifically, larval
parasitoids decreased the probability of observing egg parasitoids at
high gall midge densities (fig. \ref{fig:EggPtoid_Selection}). Note
however that this trait-fitness relationship did not translate into a
significant difference in selection (contrast \(\beta_{\text{Pref}}\)=
0.25 {[}-0.09, 0.64{]} ). This was likely because there was a tendency
(albeit not statistically signficant) for a lower probability of
egg-parasitoid survival in complex vs.~simple food webs (contrast=0.1
{[} -0.09, 0.32{]}), which masks this trait-fitness relationship. We
also observed two other interesting trends, although neither
significantly differed between treatments. Specifically, there was a
tendency for the impact of larval parasitoids to increase nonlinearly at
higher gall midge densities (fig. \ref{fig:EggPtoid_Selection}; complex
\(\alpha_{\text{Pref:Pref}}\)= -0.49 {[}-1.04, -0.13{]}; but contrast
\(\alpha_{\text{Pref:Pref}}\)= 0.46 {[}0.02, 1.07{]}). There was also a
tendency for egg parasitoids to obtain a refuge in large gall chambers
at high gall midge densities (complex \(\alpha_{\text{Pref:Diam}}\)=
0.25 {[}0.01, 0.54{]}; but contrast \(\alpha_{\text{Pref:Diam}}\)= -0.25
{[}-0.6, 0.05{]}).

@@

The generality of our results likely depends on the relative abundance
and functional differences between consumers in a community. For
example, if consumers do not differ from each other, then we do not
expect changes in food-web complexity to modify selective constraints.
Also, many consumers may be at too low of abundances to impose selection
on their resources. Rank abundance curves \citep{Preston1948} and the
disproportionate number of weak interactions in diverse communities
\citep{Paine1992} support this notion. When consumers are abundant
though, the effect of food-web complexity will depend on whether
different species impose conflicting selection pressures or select for
distinct traits. For example, parisitoids and birds impose conflicting
selection pressures on the size of galls induced by the fly
\emph{Eurosta solidaginis} \citep{Weis1985, Abrahamson1997, Start2016}.
Recent studies in this system have shown that decreases in the relative
abundance of birds, due to either small patch sizes \citep{Start2016} or
proximity to urban areas \citep{Start2018} causes a shift from neutral
to directional selection on gall size. On the other hand, different
consumers may impose selection on different traits, favoring trait
integration and increasing genetic constraints. Examples of this include
strong genetic covariances in plant resistance to different insect
herbivores \citep{Maddox1990, Wise2007, Wise2013}, although there are
also examples where these covariances are weak
\citep{Roche1997, Barbour2015}, or vary from year-to-year
\citep{Johnson2007}. We suggest that gaining predictive insight to the
evolutionary consequences of food-web disassembly requires an
understanding of the mechanisms governing the assembly of trophic
interactions.

Our results suggest that the simplification of ecological communities
may not only directly affect connected species, but also result in
indirect evolutionary effects. In our study, this indirect effect arises
from egg parasitoids being released from intraguild predation in simpler
food webs. This release occurs more on trees with high larval densities,
which could intensify selection on gall midge oviposition preference. A
growing number of experiments over the past two decades have
demonstrated the presence and potential importance of indirect
evolutionary effects that emerge in a community context
\citep{Juenger1998, Stinchcombe2001, Lankau2007, Walsh2008, Walsh2010, terHorst2010, Sahli2011, Lau2012, terHorst2015, Schiestl2018, StartInPress}.
If indirect evolutionary effects are common \citep{Walsh2013}, then
predicting evolutionary trajectories resulting from the simplification
of food webs will require evolutionary studies to explicitly account for
the ecological networks that species are embedded in.

@@
